---
title: "Matrix-oriented modelling with ROI"
date: 2018-04-28T22:20:37+02:00
draft: false
---

## A first example

In this article, we take the Knapsack problem from before and model it using the matrix based interface in [ROI](https://CRAN.R-project.org/package=ROI). `ROI` is short for the *R Optimization Infrastructure* and is an excellent family of packages to solve a variaty of optimization problems, including MILPs.

$$
\begin{equation*}
\begin{array}{ll@{}ll}
\text{max}  & \displaystyle\sum\limits_{i=1}^{n} v_{i}x_{i} & &\\
\text{subject to}& \displaystyle\sum\limits_{i=1}^{n}   w_{i}x_{i} \leq W,  & &\\
                 &                                                x_{i} \in \{0,1\}, &i=1 ,\ldots, n&
\end{array}
\end{equation*}
$$

A first step is of course to load the package and define some model parameters:

```{r}
library(ROI)
```

```{r}
n <- 10
W <- 2
v <- runif(n)
w <- runif(n)
```

Here `v` and `w` are vectors of the length `n` that can be directly passed to ROI. 

Let's build the constraint first. As the Knapsack problem only has one constraint (i.e. one row in the constraint matrix), this step is rather simple:

```{r}
constraints <- L_constraint(w, "<=", W)
```

Then we define an optimization model:

```{r}
model <- OP(objective = v, 
            constraints = constraints,
            bounds = V_bound(li = 1:n, lb = rep.int(0, n), ui = 1:n, ub = rep.int(1, n)), 
            types = rep.int("B", n), 
            maximum = TRUE)
model
```

The parameters of `OP` are self-explaining: we create an optimization problem with the objective coefficient vector `v` and our constraint from above. We further define variable bounds for all $0 \leq x_i\leq 1$ and set the type to "B", meaning they are binary variables.

Having now formulated our problem, we can pass it to one of the many available solvers. In this case we use [GLPK](https://en.wikipedia.org/wiki/GNU_Linear_Programming_Kit).

```{r}
library(ROI.plugin.glpk)
res <- ROI_solve(model, "glpk", verbose = TRUE)
res
```

`ROI` provides function to access the optimal solution (i.e. the concrete values of all $x_i$). A $1$ means, we put the object into our knapsack, a $0$ means the opposite.

```{r}
ROI::solution(res)
```

And we are done: we solved the knapsack problem using GLPK and ROI. 
